```{r Set contrast params {{model.index}}.{{contrast.index}}}
model.index <- {{model.index}}
contrast.index <- {{contrast.index}}

count.string <- c("first", "second", "third", "fourth",
                  "fifth", "sixth", "seventh", "eight", "ninth", "tenth")
```

```{r Contrast title {{indices.char}}, results='asis'}
# Get DESeqResults from the deseq2 analysis for current contrast
dds.res <- model.contrasts[[contrast.index]]

# Parse contrasts as vector
curr.contrast <- model.comparisons[[contrast.index]] # model.comparisons is set in the _DifferentialAnalysis-model.Rmd file
contrast.vec <- unlist(strsplit(curr.contrast, " "))
contrast.char <- strsplit(
        mcols(dds.res)$description[2], 
        split = ": ")[[1]][2]
```

### Comparison {{model.index}}.{{contrast.index}}: `r contrast.char`

```{r get model contrast description {{indices.char}}}
# Same annotation you'd see for the results in the R console above the
# results data.frame
contrast.description <- paste0("\t* ",
                               paste0(mcols(dds.res)$description, 
                               collapse = "\n\t* "))
```

The `r count.string[contrast.index]` comparison is `r contrast.char`. We can view the results found using DESeq2 based on the adjusted p-value (using Benjamin-Hochberg and independent filtering) less than `r padj.alpha`.

The contrast results can be described as follows,

`r paste0(contrast.description)`

#### Model results summary

```{r Deseq summary {{indices.char}}}
dds <- dds.model

# Use the results from the combined results output
dds.res <- model.contrasts[[{{contrast.index}}]]

# Summary based on combined p-adjusted results
summary.DESeqResults(dds.res, alpha = padj.alpha)
```

```{r MD Plot of result {{indices.char}}}
# Save plot file location
glimma.md.loc <- file.path(".", dds.analysis$MD.Plots[[model.index]][contrast.index])
```

```{r Glimma link {{indices.char}}, results="asis"}
cat("The results are best viewed using the MD plot using `Glimma` that shows differentially expressed in an interactive plot here\n")
cat(paste0(' <a class="btn btn-primary" href="', glimma.md.loc,'" role="button">Glimma MD plot</a>'))
```



