---
title: 'RnaSeq input data to DESeqDataSet'
author: "Lodewic van Twillert"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_caption: yes
    theme: cerulean
params:
  echo:
    value: FALSE
    input: checkbox
  dds:
    label: "DESeqDataset .RData file"
    value: NA
    input: file
  annotationdata:
    label: "Gene annotation file"
    value: NA
    input: file
bibliography: bibliography.bib
---

```{r include = FALSE}
# If you are using knit_with_params() to render this document with a little shiny app for parameter input
# You may have to reset the max requestsize in Shiny to upload files larger than 5mb.
# Run the following command in your console or other environment to allow larger than 5mb files.

# options(shiny.maxRequestSize=30*1024^2)
```


```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = params$echo)
library(DT)
library(RnaSeqAnalysis)
library(DESeq2)
# Default example data
dds.file <- ifelse(params$dds == "NA",
                   system.file("ExampleData/Liver_DDS.RData", package = "RnaSeqAnalysis"),
                   params$dds)
```

```{r Datatable function}
ShowDataTable <- function(df) {
  DT::datatable(df,
    class   = 'cell-border stripe compact hover',
    escape  = F, selection = 'multiple',
    extensions = "Scroller",
    filter = "top",
    options = list(
      scrollY = "60vh",
      scroller = TRUE,
      deferRender = TRUE,
      scrollX = TRUE,
      autoWidth  = T,
      # columnDefs = list(list(
      #   targets  = 1:ncol(df),
      #   render   = JS(
      #     "function(data, type, row, meta) {",
      #     "return type === 'display' && data.length > 50 ?",
      #     "'<span title=\"' + data + '\">' +
      #     data.substr(0, 50) + '...</span>' : data;", "}"))),
      paging = TRUE, server = T)
  )
}
```

# Input options

```{r}
knitr::kable(data.frame(Value = unlist(params)))
```

# Rna-Seq data

## DESeq2 dataset

We use a package called [DESeq2](https://bioconductor.org/packages/release/bioc/html/DESeq2.html)[@deseq2] to do differential analysis with. This package also comes with a `DESeqDataSet` format to combine count, sample data. A DESeq2 dataset also combines count and sample data in a convenient way for exploratory/clustering analysis.

```{r Load dds}
# Cache creation and computation of DESeqDataSet?
#   Estimating disperions and the VST may take a while.
load(dds.file) # Loads "dds" into the environment
vst.included <- !is.null(mcols(dds)$vst)
```

```{r Show dds}
dds
``` 

`DESeq2` also includes a favored normalization method and a **Variance Stabilizing Transformation**. The count normalization is needed to minimize the effects of the differences in sequencing depth per sample. While there are many normalization methods, here we use the `DESeq2` method. 

The Variance Stabilizing Transformation is a transformation similar to taking the *Log* of the counts, but more tailored to gene count data and also suited for zero-values. We leverage this transformation only for exploratory analysis, especially clustering analysis, but it is not used for differential analysis.

```{r Annotate rnaSeq data, eval = FALSE}
# If ANNOTATIONFILE is not specified then we use databases for annotation
# This requires the SPECIES (i.e. 'Human', 'Mouse', etc.)
# And the KEYTYPE (i.e. ENSEMBL, SYMBOL etc.)
if (try(is.na(params$annotationdata)) | params$annotationdata == "NA") {
  annotationData <- AnnotateCounts(countTable,
                                   species = params$species,
                                   key.type = params$keytype)
# Otherwise we read the data from the specified ANNOTATIONFILE
} else annotationData <- ReadAnnotation(params$annotationdata)
```

## Sample data {.tabset}

The sample data describes subject features.
We have sample data with a total of `r nrow(colData(dds))` samples with `r ncol(colData(dds))` features.
Tables are shown with the [DataTable](https://rstudio.github.io/DT/) package[@DT].

### Sample table

```{r Show column data}
# So this datatable needs a ton of options to display nicely
#   The contents are more important!
ShowDataTable(data.frame(colData(dds)))
```

### Sample feature names

`r paste0("\t* ", names(colData(dds)), collapse = "\n")`

## Expression data {.tabset}

Only showing first 100 genes for document size.
The count data consists of `r nrow(dds)` genes with non-zero expression of `r ncol(dds)` samples.

Raw counts must be non-negative integer values, while the Normalized counts are
normalized with `DESeq2` using a form of quantile normalization.

### Raw counts

```{r}
ShowDataTable(data.frame(counts(dds))[1:100,])
```

### Normalized counts

```{r}
ShowDataTable(data.frame(counts(dds, normalized = TRUE))[1:100,])
```


# References



