---
title: 'RnaSeq: Read input data to DESeqDataSet'
author: "Lodewic van Twillert"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_caption: yes
    theme: cerulean
    self_contained: true
params:
  echo:
    value: FALSE
    input: checkbox
  coldata:
    label: "Coldata file"
    value: "InputFiles/HEEL_ColData.csv"
    input: file
  countdata:
    label: "Count file"
    value: "InputFiles/HEEL_Counts.csv"
    input: file
  annotationdata:
    label: "Gene annotation file"
    value: NA
    input: file
  species:
    label: "Gene species"
    value: "Mouse"
    input: select
    choices: [Human, Mouse, Rat]
  keytype:
    label: "Gene mapping keytype"
    value: "ENSEMBL"
    input: select
    choices: [ENSEMBL, SYMBOL, Entrez]
  outputdir:
    label: "Sub-directory of this document for output files"
    value: "ReportOutput/"
  output:
    label: "Output .Rdata file"
    value: "ReadDataOutput.RData"
---

```{r include = FALSE}
# If you are using knit_with_params() to render this document with a little shiny app for parameter input
# You may have to reset the max requestsize in Shiny to upload files larger than 5mb.
# Run the following command in your console or other environment to allow larger than 5mb files.

# options(shiny.maxRequestSize=30*1024^2)
```


```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = params$echo)
library(DT)
library(RnaSeqAnalysis)
library(DESeq2)
```

# Input options

```{r}
knitr::kable(data.frame(Value = unlist(params)))
```

# Input data

```{r Read input data}
# Use general functions to read rnaSeq data
# The separator is determined programmatically
#   CURRENTLY ONLY FOR .CSV FILES
countTable <- ReadCounts(params$countdata, sep = NULL, sort.samples = TRUE)
colData <- ReadColData(params$coldata, sep = NULL, sort.samples = TRUE)
```

```{r Annotate rnaSeq data}
# If ANNOTATIONFILE is not specified then we use databases for annotation
# This requires the SPECIES (i.e. 'Human', 'Mouse', etc.)
# And the KEYTYPE (i.e. ENSEMBL, SYMBOL etc.)
if (try(is.na(params$annotationdata)) | params$annotationdata == "NA") {
  annotationData <- AnnotateCounts(countTable,
                                   species = params$species,
                                   key.type = params$keytype)
# Otherwise we read the data from the specified ANNOTATIONFILE
} else annotationData <- ReadAnnotation(params$annotationdata)
```

```{r Create DESeqDataSet}
dds <- DESeqDataSetFromMatrix(countTable, colData, design = ~1)
dds <- estimateSizeFactors(dds)
dds <- estimateDispersions(dds)
```

### Sample data

```{r Show column data}
DT::datatable(data.frame(colData(dds)), options = list(scrollX = TRUE))
```

Display of **sample data** with a total of `r nrow(colData)` samples with `r ncol(colData)` features. The samples features are named,

`r paste0("\t* ", names(colData), collapse = "\n")`

### Expression data {.tabset}

Only showing first 100 genes for document size.
The count data consists of `r nrow(countTable)` genes with non-zero expression of `r ncol(countTable)` samples.
The **`r params$species`** genes are mapped to **`r params$keytype`** names.

#### Raw counts

```{r}
DT::datatable(data.frame(counts(dds))[1:100,], options = list(scrollX = TRUE))
```

#### Normalized counts

```{r}
DT::datatable(data.frame(counts(dds, normalized = TRUE))[1:100,], options = list(scrollX = TRUE))
```


# Output data

Output data to **`r file.path(params$outputdir, params$output)`**

```{r}
if (!dir.exists(params$outputdir)) dir.create(params$outputdir)
save(dds, file = file.path(params$outputdir, params$output))
```

